---
title: "Randomization with Wrapping"
author: "David Ye"
date: "`r Sys.Date()`"
output: html_document
---
```{r}
library(truncnorm)
```

Method 0:
randomly shuffle every point
```{r}
shuffle_t <- function(object, tone = FALSE) {
  neurons <- length(object[,1,1])
  trials <- length(object[1,,1])
  all_tests <- matrix(NA, nrow = neurons, ncol = trials)
  # randomization distribution with n = 1000 samples
  n <- 1000
  rand_dist <- rep(NA, n)
  pre_window_tone <- 100:300
  post_window_tone <- 301:400
  pre_window_shock_food <- 301:429
  post_window_shock_food <- 430:550
  for (i in 1:neurons) {
    for (j in 1:trials) {
      if(tone) {
        combined <- object[i,j,c(pre_window_tone,post_window_tone)]
        for (k in 1:n) {
          shuffled <- sample(combined)
          pre_shuffle <- shuffled[1:length(pre_window_tone)]
          post_shuffle <- shuffled[length(pre_window_tone)+1:length(shuffled)]
          rand_dist[k] <- as.numeric((t.test(x = post_shuffle,y = pre_shuffle)$statistic))
        }
        #plot(hist(rand_dist))
        obs_stat <- as.numeric((t.test(x = object[i,j,post_window_tone],y = object[i,j,pre_window_tone])$statistic))
      }
      else {
        combined <- object[i,j,c(pre_window_shock_food,post_window_shock_food)]
        for (k in 1:n) {
          shuffled <- sample(combined)
          pre_shuffle <- shuffled[1:length(pre_window_shock_food)]
          post_shuffle <- shuffled[length(pre_window_shock_food)+1:length(shuffled)]
          rand_dist[k] <- as.numeric((t.test(x = post_shuffle,y = pre_shuffle)$statistic))
        }
        #plot(hist(rand_dist))
        obs_stat <- as.numeric((t.test(x = object[i,j,post_window_shock_food],y = object[i,j,pre_window_shock_food])$statistic))
      }
      # lower p_higher means more significant increase in test statistic
      # lower p_lower means more significant decrease in test statistic
      p_higher <- mean(rand_dist >= obs_stat)
      p_lower <- mean(rand_dist <= obs_stat)
      # plot(hist(rand_dist)) # for null distribution visualization
      # two-tailed thus multiply smaller p-value by 2
      all_tests[i,j] <- 2*(min(p_higher,p_lower))
    }
  }
  return (all_tests)
}
```

Method 1:
randomly wrap uniformly between 1 and T

Maybe print out distribution of statistic and compare distribution to 
Method 1 etc.
```{r}
unif_wrap_t <- function(object, tone = FALSE) {
  neurons <- length(object[,1,1])
  trials <- length(object[1,,1])
  all_tests <- matrix(NA, nrow = neurons, ncol = trials)
  # randomization distribution with n = 1000 samples
  n <- 1000
  rand_dist <- rep(NA, n)
  pre_window_tone <- 100:300
  post_window_tone <- 301:400
  pre_window_shock_food <- 301:429
  post_window_shock_food <- 430:550
  for (i in 1:neurons) {
    for (j in 1:trials) {
      if(tone) {
        combined <- object[i,j,c(pre_window_tone,post_window_tone)]
        wrap_around <- round(runif(n, min=1, max=length(combined)), digits = 0)
        # print(wrap_around)
        for (k in 1:n) {
          # avoid combined[1:0] issue --> means stay as original
          if (wrap_around[k] != 1) {
            wrapped <- c(combined[wrap_around[k]:length(combined)], 
                         combined[1:wrap_around[k] - 1])
          }
          else {
            wrapped <- combined
          }
          pre_wrap <- wrapped[1:length(pre_window_tone)]
          post_wrap <- wrapped[length(pre_window_tone)+1:length(wrapped)]
          #plot(neuron_activity_portrait_marked(wrapped))
          rand_dist[k] <- as.numeric((t.test(x = post_wrap,
                                             y = pre_wrap)$statistic))
        }
        obs_stat <- as.numeric((t.test(x = object[i,j,post_window_tone],
                                       y = object[i,j,pre_window_tone])$statistic))
      }
      # food/shock
      else {
        combined <- object[i,j,c(pre_window_shock_food,post_window_shock_food)]
        # plot(neuron_activity_portrait(combined))
        wrap_around <- round(runif(n, min=1, max=length(combined)), digits = 0)
        for (k in 1:n) {
          if (wrap_around[k] != 1) {
            wrapped <- c(combined[wrap_around[k]:length(combined)], 
                         combined[1:wrap_around[k] - 1])
          }
          else {
            wrapped <- combined
          }
          pre_wrap <- wrapped[1:length(pre_window_shock_food)]
          post_wrap <- wrapped[length(pre_window_shock_food)+1:length(wrapped)]
          #plot(neuron_activity_portrait_marked(wrapped))
          rand_dist[k] <- as.numeric((t.test(x = post_wrap,
                                             y = pre_wrap)$statistic))
        }
        obs_stat <- as.numeric((t.test(x = object[i,j,post_window_shock_food],
                                       y = object[i,j,pre_window_shock_food])$statistic))
      }
      # lower p_higher means more significant increase in test statistic
      # lower p_lower means more significant decrease in test statistic
      p_higher <- mean(rand_dist >= obs_stat)
      p_lower <- mean(rand_dist <= obs_stat)
      # two-tailed thus multiply smaller p-value by 2
      all_tests[i,j] <- 2*(min(p_higher,p_lower))
    }
  }
  return (all_tests)
}
```

Method 2:
Random wrapping with Gaussian centered at middle of time series and
standard deviation of 50 for tone and 40 for food/shock (arbitrarily chosen) with max min cutoff
```{r}
gauss_wrap_t <- function(object, tone = FALSE) {
  neurons <- length(object[,1,1])
  trials <- length(object[1,,1])
  all_tests <- matrix(NA, nrow = neurons, ncol = trials)
  # randomization distribution with n = 1000 samples
  n <- 1000
  rand_dist <- rep(NA, n)
  for (i in 1:neurons) {
    for (j in 1:trials) {
      if(tone) {
        combined <- object[i,j,100:400]
        for (k in 1:n) {
          wrap_around <- round(rtruncnorm(n=1, a=1, b=301, mean=301/2, sd=50),
                               digits = 0)
          wrapped <- c(combined[wrap_around:length(combined)], combined[1:wrap_around - 1])
          rand_dist[k] <- as.numeric((t.test(x = wrapped[202:301],y = wrapped[1:201],
                                    conf.level = 0.95)$statistic))
        }
        obs_stat <- as.numeric((t.test(x = object[i,j,301:400],y = object[i,j,100:300],
                                     conf.level = 0.95)$statistic))
      }
      # food/shock
      else {
        combined <- object[i,j,301:550]
        for (k in 1:n) {
          wrap_around <- round(rtruncnorm(n=1, a=1, b=250, mean=250/2, sd=40),
                               digits = 0)
          wrapped <- c(combined[wrap_around:length(combined)], combined[1:wrap_around - 1])
          rand_dist[k] <- as.numeric((t.test(x = wrapped[130:250],y = wrapped[1:129],
                                    conf.level = 0.95)$statistic))
        }
        obs_stat <- as.numeric((t.test(x = object[i,j,430:550],y = object[i,j,301:429],
                                     conf.level = 0.95)$statistic))
      }
      # lower p_higher means more significant increase in test statistic
      # lower p_lower means more significant decrease in test statistic
      p_higher <- sum(rand_dist >= obs_stat)/(n)
      p_lower <- sum(rand_dist <= obs_stat)/(n)
      # two-tailed thus multiply smaller p-value by 2
      all_tests[i,j] <- 2*(min(p_higher,p_lower))
    }
  }
  return (all_tests)
}

```



Method 3:
K-Chunking with 8 Chunks
```{r}
k_chunk_t <- function(object, tone = FALSE) {
  neurons <- length(object[,1,1])
  trials <- length(object[1,,1])
  all_tests <- matrix(NA, nrow = neurons, ncol = trials)
  n <- 1000
  num_chunks <- 8
  rand_dist <- rep(NA, n)
  pre_window_tone <- 100:300
  post_window_tone <- 301:400
  pre_window_shock_food <- 301:429
  post_window_shock_food <- 430:550
  for (i in 1:neurons) {
    for (j in 1:trials) {
      if(tone) {
        combined <- object[i,j,c(pre_window_tone,post_window_tone)]
        # plot
        # plot(neuron_activity_portrait(combined))
        chunk_size <- round(length(combined)/num_chunks, digits = 0)
        chunk <- split(combined, ceiling(seq_along(combined)/chunk_size))
        for (k in 1:n) {
          wrapped <- sample(chunk)
          wrapped <- unlist(wrapped)
          # plot
          # plot(neuron_activity_portrait_marked(wrapped))
          pre_wrap <- wrapped[1:length(pre_window_tone)]
          post_wrap <- wrapped[length(pre_window_tone)+1:length(wrapped)]
          rand_dist[k] <- as.numeric((t.test(x = post_wrap,
                                             y = pre_wrap)$statistic))
        }
        #plot(hist(rand_dist))
        obs_stat <- as.numeric((t.test(x = object[i,j,post_window_tone],y = object[i,j,pre_window_tone])$statistic))
      }
      # food/shock
      else {
        combined <- object[i,j,c(pre_window_shock_food,post_window_shock_food)]
        # plot
        # plot(neuron_activity_portrait(combined))
        chunk_size <- round(length(combined)/num_chunks, digits = 0)
        chunk <- split(combined, ceiling(seq_along(combined)/chunk_size))
        for (k in 1:n) {
          wrapped <- sample(chunk)
          wrapped <- unlist(wrapped)
          # plot
          # plot(neuron_activity_portrait_marked(wrapped))
          pre_wrap <- wrapped[1:length(pre_window_tone)]
          post_wrap <- wrapped[length(pre_window_tone)+1:length(wrapped)]
          rand_dist[k] <- as.numeric((t.test(x = post_wrap,
                                             y = pre_wrap)$statistic))
        }
        #plot(hist(rand_dist))
        obs_stat <- as.numeric((t.test(x = object[i,j,post_window_shock_food],
                                       y = object[i,j,pre_window_shock_food])$statistic))
      }
      # lower p_higher means more significant increase in test statistic
      # lower p_lower means more significant decrease in test statistic
      p_higher <- sum(rand_dist >= obs_stat)/(n)
      p_lower <- sum(rand_dist <= obs_stat)/(n)
      # two-tailed thus multiply smaller p-value by 2
      all_tests[i,j] <- 2*(min(p_higher,p_lower))
    }
  }
  return (all_tests)
}
```

```{r}
load("3D_Activation_Object.RData")
Food_Day_10 <- Food_Day_10[1:5,,]
#Shock_Day_11 <- Shock_Day_11[1:2,,]
#Test_Day_12 <- Test_Day_12[1:2,,]
```

```{r}
day10_tone_t_ <- all_t_test(Food_Day_10,tone = TRUE)
day10_food_t_ <- all_t_test(Food_Day_10,tone = FALSE)
day11_tone_t_ <- all_t_test(Shock_Day_11,tone = TRUE)
day11_shock_t_ <- all_t_test(Shock_Day_11,tone = FALSE)
day12_ <- all_t_test(Test_Day_12,tone = TRUE)
#save(day10_tone_t, day10_food_t, day11_tone_t, day11_shock_t, day12,file = "t_test.RData")

day10_tone_MW <- all_Mann_Whitney_test(Food_Day_10,tone = TRUE)
day10_food_MW <- all_Mann_Whitney_test(Food_Day_10,tone = FALSE)
day11_tone_MW <- all_Mann_Whitney_test(Shock_Day_11,tone = TRUE)
day11_shock_MW <- all_Mann_Whitney_test(Shock_Day_11,tone = FALSE)
day12_tone_MW <- all_Mann_Whitney_test(Test_Day_12,tone = TRUE)
#save(day10_tone_MW, day10_food_MW, day11_tone_MW, day11_shock_MW, day12_tone_MW,file = "MW_test.RData")
```

```{r}
day10_Food_shuffle_t <- shuffle_t(Food_Day_10, tone = FALSE)
day10_Tone_shuffle_t <- shuffle_t(Food_Day_10, tone = TRUE)
day11_Shock_shuffle_t <- shuffle_t(Shock_Day_11, tone = FALSE)
day11_Tone_shuffle_t <- shuffle_t(Shock_Day_11, tone = TRUE)
day12_shuffle_t <- shuffle_t(Test_Day_12, tone = TRUE)
#save(day10_Food_shuffle_t, day10_Tone_shuffle_t, day11_Shock_shuffle_t, day11_Tone_shuffle_t, day12_shuffle_t,file = "shuffle_t.RData")
```

```{r}

day10_Food_unifwrap_t <- unif_wrap_t(Food_Day_10, tone = FALSE)
day10_Tone_unifwrap_t <- unif_wrap_t(Food_Day_10, tone = TRUE)
day11_Shock_unifwrap_t <- unif_wrap_t(Shock_Day_11, tone = FALSE)
day11_Tone_unifwrap_t <- unif_wrap_t(Shock_Day_11, tone = TRUE)
day12_unifwrap_t <- unif_wrap_t(Test_Day_12, tone = TRUE)
#save(day10_Food_unifwrap_t, day10_Tone_unifwrap_t, day11_Shock_unifwrap_t, day11_Tone_unifwrap_t, day12_unifwrap_t,file = "unifwrap_t.RData")
```

```{r}
day10_Food_gausswrap_t <- gauss_wrap_t(Food_Day_10, tone = FALSE)
day10_Tone_gausswrap_t <- gauss_wrap_t(Food_Day_10, tone = TRUE)
day11_Shock_gausswrap_t <- gauss_wrap_t(Shock_Day_11, tone = FALSE)
day11_Tone_gausswrap_t <- gauss_wrap_t(Shock_Day_11, tone = TRUE)
day12_gausswrap_t <- gauss_wrap_t(Test_Day_12, tone = TRUE)
#save(day10_Food_gausswrap_t, day10_Tone_gausswrap_t, day11_Shock_gausswrap_t, day11_Tone_gausswrap_t, day12_gausswrap_t,file = "gausswrap_t.RData")
```

```{r}
day10_Food_kchunk_t <- k_chunk_t(Food_Day_10, tone = FALSE)
day10_Tone_kchunk_t <- k_chunk_t(Food_Day_10, tone = TRUE)
day11_Shock_kchunk_t <- k_chunk_t(Shock_Day_11, tone = FALSE)
day11_Tone_kchunk_t <- k_chunk_t(Shock_Day_11, tone = TRUE)
day12_kchunk_t <- k_chunk_t(Test_Day_12, tone = TRUE)
save(day10_Food_kchunk_t, day10_Tone_kchunk_t, day11_Shock_kchunk_t, day11_Tone_kchunk_t, day12_kchunk_t,file = "kchunk_t.RData")
```
